An Abstraction-based Approach to Network Synthesis

### 2016-11-03 paper/abstraction/paper.tex
background
adding compilation to abgp

For each regular expression $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. An automaton for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma = G^A$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The Product Graph is a tuple ($V'$, $E'$, $s$, $P$) with
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of path ranks.

The Product Graph is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$ is a valid topology link.
%
We add edges from the start node $s$ to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the Product Graph ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$. 
%
We write $worse(m,n)$ if $P(n) = \emptyset$ or $\min ~ P(m) \geq \min ~ P(n)$, which means that paths ending at PGIR node $m$ are worse (higher rank) than paths ending at $n$.
%
We write $m \approx n$ whenever two vertices $m,n = (l, \dots) \in V'$ share the same topology location $l$.

@@

@@

The first idea is to tag every protocol message with a BGP community value that relays the state of each automaton when advertising a route to a neighbor. Routers can then allow route advertisements that correspond to an edge in the Product Graph (\IE, from a neighbor with the correct state) and block those that don't. This ensures that any path computed by BGP will be a valid path in the Product Graph, and thus is allowed by the policy. For example, in the concrete Product Graph in figure~\ref{fig:example-compilation}, router $S1$ will allow a route advertisement from its neighbor $P1$ and also from neighbor $P2$. If it uses neighbor $P1$, then it will add the tag $(1,0)$ before informing its neighbors about its route. On the other hand, if it uses neighbor $P2$, then it will add the tag $(0,1)$ before informing its neighbors about the route.

%The first step is to generate per-device BGP import and export filters based on the edges in the Product Graph. The idea is to attach a tag to every route advertisement that routers send to their neighbors that encodes the current state of the automata. For example... \ryan{show example here}. This ensures that BGP will only find paths that exist in the Product Graph, and are thus allowed by the policy.

However, letting BGP find some allowed path is not enough. \sysname must ensure BGP always computes the \emph{best} path according to the user's preferences, even when failures occur. The second idea is to totally order advertisements each router might receive from neighbors in the Product Graph by preference. For example, to satisfy the policy that leaving through $\sf{P1}$ is preferred to leaving through $\sf{P2}$, router $S1$ should prefer a route it hears from $P1$ over one from $P2$. Similarly, router $A1$ should prefer routes tagged with $(1,0)$ over those tagged with $(0,1)$.

@@

\begin{defn}
The product graph can be viewed as a labelled transition system where
$m\overset{l}{\rightarrow}n$ if edge $(m,n) \in E'$ and $loc(n) = l$. We write $m \leq m'$ if $m$ simulates $m'$ with respect to the ``worse" relation on node rankings.
\end{defn}

@@

For example, in the concrete product graph we have $(S1,0,1) \leq (S1,1,0)$ but not the other way around. Therefore, advertisements received at $S1$ in state $(0,1)$ must be preferred to those received in state $(1,0)$. Observe also that in the abstract product graph, node $(S,0,1) \leq (S,1,0)$. This leads us the the following lemma:

\begin{lem}
\label{lem:preference}
$m \leq m'$ in the concrete product graph iff $f(m) \leq f(m')$ in the abstract product graph under the lifted graph homomorphism $f$.
\end{lem}

Intuitively, Lemma~\ref{lem:preference} hints that we will be able to infer BGP local preferences over the abstract product graph rather than the concrete product graph.

\subsection{Substitution Theorem}

\begin{easylist}[itemize]
& First show how to compile to the Product Graph, and from there to the ABGP
& Show/Define safety analysis simulation relation for both concrete and abstract product graphs
  and how it holds for both.
& Next define the concretization & subst. functions for translating the abstract policy
   into the concrete policy, using the ABGP example for reference
& Lemma: if safety analysis holdls in the abstract graph, it holds for the concrete one,
   and the preferences are consistent.
& Give abstraction theorem: commuting of concretization and compilation
\end{easylist}

>>>>

For each regular expression $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. An automaton for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma = G^A$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The Product Graph is a tuple ($G'$, $s$, $P$) where $G' = (V',E')$ is a new graph with 
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $s$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$ , which maps nodes in the product graph to a set of path ranks.

The Product Graph is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$ is a valid topology link.
%
We add edges from the start node $s$ to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the Product Graph ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\tilde{m} = l$ to extract the topology location from a product graph node, when $m = (l, \dots) \in V'$.

@@

From Figure~\ref{fig:example-compilation} we can see that both the concrete and abstract product graphs share a similar structure. This leads us to the following observation

\begin{defn}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete product graph $PG = (G',s,P)$ and abstract product graph $PG^A = (G'^A, s^A, P^A)$, then we can construct a new, lifted homomorphism $f_{pg} : G' \rightarrow G'^A$ in the following way: 
\[ \begin{array}{rcl}
  f_{pg}( s ) & = & s^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}

@@

\para{Route filtering}

The first idea is to tag every protocol message with a BGP community value that relays the state of each automaton when advertising a route to a neighbor. Routers can then allow route advertisements that correspond to an edge in the Product Graph (\IE, from a neighbor with the correct state) and block those that don't. This ensures that any path computed by BGP will be a valid path in the Product Graph, and thus is allowed by the policy. For example, in the concrete Product Graph in figure~\ref{fig:example-compilation}, router $S1$ will allow a route advertisement from its neighbor $P1$ and also from neighbor $P2$. If it uses neighbor $P1$, then it will add the tag $(1,0)$ before informing its neighbors about its route. On the other hand, if it uses neighbor $P2$, then it will add the tag $(0,1)$ before informing its neighbors about the route.
%
However, letting BGP find some allowed path is not enough. \sysname must ensure BGP always computes the \emph{best} path according to the user's preferences, even when failures occur.

\para{Preference Search}
The second idea is to totally order advertisements each router might receive from neighbors in the Product Graph by preference. For example, to satisfy the policy that leaving through $\sf{P1}$ is preferred to leaving through $\sf{P2}$, router $S1$ should prefer a route it hears from $P1$ over one from $P2$. Similarly, router $A1$ should prefer routes tagged with $(1,0)$ over those tagged with $(0,1)$.

@@

\begin{defn}
Let $m \geq_{rank} n$ be a relation over Product Graph vertices that holds iff $P(n) = \emptyset$ or $\min P(m) \geq \min P(n)$. Intuitively $m \geq_{rank} n$ means that paths ending at PGIR node $m$ are no better (higher rank) than paths ending at $n$.
%
\end{defn}
\noindent
%

\begin{defn}
The product graph can be viewed as a labelled transition system by pushing the topology location from each directed edge's target node onto the edge. That is, $m\overset{l}{\rightarrow}n$ if edge $(m,n) \in E'$ and $\tilde{n} = l$. 
\end{defn}

For example, in concrete product graph we would have the transition $(S1,1,0)\overset{A2}{\rightarrow}(A2,1,0)$

\begin{defn}
We write $m \leq m'$ if the subgraph reachable from $m$ and $m'$ respectively form a simulation relation with respect to the $\geq_{rank}$ relation on node rankings.
\end{defn}

@@

For example, in the concrete product graph we have $(S1,0,1) \leq (S1,1,0)$ since nodes on the left side of the product graph can always match transitions made on the right hand side while always maintaining a node ranking that is at least as good as that on the right hand side (since they match the preferred rank 1 automaton). This relation does not hold the other way around. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received in state $(0,1)$.

Notice also that in the abstract product graph, node $(S,0,1) \leq (S,1,0)$. This leads us the the following observation:

\begin{lem}
\label{lem:preference}
$m \leq m'$ in the concrete product graph iff $f_{pg}(m) \leq f_{pg}(m')$ in the abstract product graph under the lifted graph homomorphism $f_{pg}$.
\end{lem}

Lemma~\ref{lem:preference} gives us reason to believe that we can perform compilation over the abstract product graph and ensure that the inferred device-level preferences will be correct when instantiating the template with a concrete network.

\newcommand{\highlight}[1]{%
  \colorbox{red!50}{$\displaystyle#1$}}
\newcommand{\Router}[1]{\KW{Router} #1:}
\newcommand{\REGEX}[1]{\texttt{regex}(#1)}
\newcommand{\PEER}{\texttt{peer}}
\newcommand{\PREFIX}{\texttt{prefix}}
\newcommand{\IF}{\texttt{if}}
\newcommand{\THEN}{\texttt{then}}
\newcommand{\COMM}{\texttt{comm}}
\newcommand{\MED}{\texttt{MED}}
\newcommand{\Arrow}{\ensuremath{\leftarrow}}

\para{BGP configurations}

\begin{figure*}[t!]
  \begin{minipage}[t]{.54\linewidth}
    \hdr{Concrete Configuration}{}
    \begin{code}
      \Router{S1}
        Match[100] \PREFIX=GP1, \PEER=*
          \PEER \Arrow *
        Match[100] \PREFIX=GP2, \PEER=*
          \PEER \Arrow *
        ...
        Match[100] \PREFIX=LP1, \PEER=\{AG1, ..., AL1, ...\}
          \PEER \Arrow \{AG1, ..., AL1, ...\}
        Match[100] \PREFIX=LP2, \PEER=\{AG1, ..., AL1, ...\}
          \PEER \Arrow \{AG1, ..., AL1, ...\}
        ...
        Match[110] \PREFIX=0.0.0.0/[0..32], \PEER=P1, 
          \PEER \Arrow *, \COMM \Arrow (1,0)
        Match[100] \PREFIX=0.0.0.0/[0..32], \PEER=P2, 
          \PEER \Arrow *, \COMM \Arrow (0,1)
        ...
      \end{code}
  \end{minipage}
  %
  \begin{minipage}[t]{.4\linewidth}
    \hdr{Template Configuration}{}
    \begin{code}
      \Router{S}
        Match[100] \PREFIX=\$GP, \PEER=*
          \PEER \Arrow *
        Match[100] \PREFIX=\$LP, \PEER=\{AG, AL\}
          \PEER \Arrow \{AG, AL\}
        Match[110] 0.0.0.0/[0..32], \PEER=P1, 
          \PEER \Arrow *, \COMM \Arrow (1,0)
        Match[100] 0.0.0.0/[0..32], \PEER=P2, 
          \PEER \Arrow *, \COMM \Arrow (0,1)
        ...
      \end{code}
  \end{minipage}%

  \caption{Spine configurations for concrete and abstract policies.}
  \label{fig:bgp-configs}
\end{figure*}

By combining the tagging, filtering and preference mechanisms, we can generate safe BGP configurations. In general, configurations are generated by combining the per-prefix configurations in the same order specified in the policy. Figure~\ref{fig:bgp-configs} shows partial a configuration for the spine routers in the concrete and abstract policies.

\subsection{Compilation Correctness}

\begin{easylist}[itemize]
& First show how to compile to the Product Graph, and from there to the ABGP
& Next define the concretization & subst. functions for translating the abstract policy
   into the concrete policy, using the ABGP example for reference
& Give abstraction theorem: commuting of concretization and compilation
\end{easylist}
